Supporting information - Bayesian regression facilitates quantitative modelling of cell metabolism

### 2024-01-09 paper/main.qmd
editorial changes

---
title: "Bayesian regression facilitates quantitative modelling of cell metabolism"
keywords: Bayesian inference, kinetic models of cell metabolism, 
          multiomics integration, ordinary differential equations,
          regulatory anlaysis
abstract: >
  This paper presents Maud, a command-line application that implements Bayesian
  statistical inference for kinetic models of biochemical metabolic reaction
  networks. Maud takes into account quantitative information from omics
  experiments and background knowledge, as well as structural information about
  kinetic mechanisms, regulatory interactions and enzyme knockouts. Our paper
  reviews the existing options in this area, presents a case study illustrating
  how Maud can be used to analyse a metabolic network and explains the
  biological, statistical and computational design decisions underpinning Maud.
number-sections: true
author: 
  - name: "Teddy Groves"
    orcid: 0000-0002-7109-3270
    email: tedgro@biosustain.dtu.dk
    affiliations: 
      - id: biosustain
        name: DTU
        department: The Novo Nordisk Foundation Center for Biosustainability
        city: Kongens Lyngby
        country: Denmark
    attributes:
      corresponding: true
      equal-contributor: true
    note: These authors contributed equally to this research.
  - name: "Nicholas Luke Cowie"
    orcid: 0000-0003-3367-2105
    affiliations: 
      - ref: biosustain
    attributes:
      equal-contributor: true
    note: These authors contributed equally to this research.
  - name: "Lars Keld Nielsen"
    orcid: 0000-0001-8191-3511
    affiliations: 
      - id: uq
        name: The University of Queensland
        department: Australian Institute for Bioengineering and Nanotechnology (AIBN)
        city: St Lucia
        country: Australia
      - ref: biosustain
bibliography: bibliography.bib
format:
  acs-pdf:
    keep-tex: true
    journal:
      id: asbcd6  #This is an identifier for the target journal: from https://pubs.acs.org/, the identifier is the part of the URL after https://pubs.acs.org/journal/, e.g. for J. Org. Chem. it would be joceah.
      type: article  #article, note, communication, review, letter, or perspectivee
      layout: traditional #traditional or twocolumn, twocolumn currently broken for longtables
      hide-email: false # if enabled, hides email addresses (and fax/phone)
  acs-html:
   include-in-header: mathjax.html
   embed-resources: true
  docx: default
---

@@

# References

>>>>

---
title: "Bayesian regression facilitates quantitative modelling of cell metabolism"
keywords: Bayesian inference, kinetic models of cell metabolism, 
          multiomics integration, ordinary differential equations,
          regulatory anlaysis
abstract: >
  This paper presents Maud, a command-line application that implements Bayesian
  statistical inference for kinetic models of biochemical metabolic reaction
  networks. Maud takes into account quantitative information from omics
  experiments and background knowledge, as well as structural information about
  kinetic mechanisms, regulatory interactions and enzyme knockouts. Our paper
  reviews the existing options in this area, presents a case study illustrating
  how Maud can be used to analyse a metabolic network and explains the
  biological, statistical and computational design decisions underpinning Maud.
number-sections: true
author: 
  - name: "Teddy Groves"
    orcid: 0000-0002-7109-3270
    email: tedgro@biosustain.dtu.dk
    affiliations: 
      - id: biosustain
        name: DTU
        department: The Novo Nordisk Foundation Center for Biosustainability
        city: Kongens Lyngby 2800
        country: Denmark
    attributes:
      corresponding: true
      equal-contributor: true
    note: These authors contributed equally to this research.
  - name: "Nicholas Luke Cowie"
    orcid: 0000-0003-3367-2105
    affiliations: 
      - ref: biosustain
    attributes:
      equal-contributor: true
    note: These authors contributed equally to this research.
  - name: "Lars Keld Nielsen"
    orcid: 0000-0001-8191-3511
    affiliations: 
      - id: uq
        name: The University of Queensland
        department: Australian Institute for Bioengineering and Nanotechnology (AIBN)
        city: St Lucia 4067
        country: Australia
      - ref: biosustain
bibliography: bibliography.bib
format:
  acs-pdf:
    keep-tex: true
    journal:
      id: asbcd6  #This is an identifier for the target journal: from https://pubs.acs.org/, the identifier is the part of the URL after https://pubs.acs.org/journal/, e.g. for J. Org. Chem. it would be joceah.
      type: article  #article, note, communication, review, letter, or perspectivee
      layout: traditional #traditional or twocolumn, twocolumn currently broken for longtables
      hide-email: false # if enabled, hides email addresses (and fax/phone)
  acs-html:
   include-in-header: mathjax.html
   embed-resources: true
  docx: default
---

@@

# Author information

## Author contribution

Teddy Groves and Nicholas Cowie: Writing code and documentation, designing
statistical model, biological model, software architecture and user interface,
writing and maintaining code, documentation and manuscript.

Lars Keld Nielsen: Supervision, writing manuscript, designing statistical model,
biological model, software architecture and user interface.   # References

## Conflict of interest

The authors declare no competing financial interest

### 2024-01-09 paper/main.tex
editorial changes

\author{Teddy Groves}
\affiliation{The Novo Nordisk Foundation Center for
Biosustainability, DTU, Kongens Lyngby, Denmark}

@@

\email{tedgro@biosustain.dtu.dk}
\author{Nicholas Luke Cowie}
\affiliation{The Novo Nordisk Foundation Center for
Biosustainability, DTU, Kongens Lyngby, Denmark}

@@

\author{Lars Keld Nielsen}
\affiliation{The Novo Nordisk Foundation Center for
Biosustainability, DTU, Kongens Lyngby, Denmark}
\affiliation{Australian Institute for Bioengineering and Nanotechnology
(AIBN), The University of Queensland, St Lucia, Australia}

@@

\title[]{Bayesian regression facilitates quantitative modelling of cell
metabolism}
\makeatletter
\begin{document}
\maketitle
\begin{abstract}
This paper presents Maud, a command-line application that implements
Bayesian statistical inference for kinetic models of biochemical
metabolic reaction networks. Maud takes into account quantitative
information from omics experiments and background knowledge, as well as
structural information about kinetic mechanisms, regulatory interactions
and enzyme knockouts. Our paper reviews the existing options in this
area, presents a case study illustrating how Maud can be used to analyse
a metabolic network and explains the biological, statistical and
computational design decisions underpinning Maud.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, breakable, interior hidden, enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi

@@

\hypertarget{references}{%
\section{References}\label{references}}

>>>>

\author{Teddy Groves}
\affiliation{The Novo Nordisk Foundation Center for
Biosustainability, DTU, Kongens Lyngby 2800, Denmark}

@@

\email{tedgro@biosustain.dtu.dk}
\author{Nicholas Luke Cowie}
\affiliation{The Novo Nordisk Foundation Center for
Biosustainability, DTU, Kongens Lyngby 2800, Denmark}

@@

\author{Lars Keld Nielsen}
\affiliation{The Novo Nordisk Foundation Center for
Biosustainability, DTU, Kongens Lyngby 2800, Denmark}
\affiliation{Australian Institute for Bioengineering and Nanotechnology
(AIBN), The University of Queensland, St Lucia 4067, Australia}

@@

\title[]{Bayesian regression facilitates quantitative modelling of cell
metabolism}
\makeatletter
\begin{document}
\maketitle
\begin{abstract}
This paper presents Maud, a command-line application that implements
Bayesian statistical inference for kinetic models of biochemical
metabolic reaction networks. Maud takes into account quantitative
information from omics experiments and background knowledge, as well as
structural information about kinetic mechanisms, regulatory interactions
and enzyme knockouts. Our paper reviews the existing options in this
area, presents a case study illustrating how Maud can be used to analyse
a metabolic network and explains the biological, statistical and
computational design decisions underpinning Maud.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, boxrule=0pt, breakable, interior hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, sharp corners]}{\end{tcolorbox}}\fi

@@

\hypertarget{author-information}{%
\section{Author information}\label{author-information}}

\hypertarget{author-contribution}{%
\subsection{Author contribution}\label{author-contribution}}

Teddy Groves and Nicholas Cowie: Writing code and documentation,
designing statistical model, biological model, software architecture and
user interface, writing and maintaining code, documentation and
manuscript.

Lars Keld Nielsen: Supervision, writing manuscript, designing
statistical model, biological model, software architecture and user
interface. \# References

\hypertarget{conflict-of-interest}{%
\subsection{Conflict of interest}\label{conflict-of-interest}}

The authors declare no competing financial interest

### 2024-01-09 paper/supporting_information.qmd
editorial changes

---
title: Supporting information
bibliography: bibliography.bib
number-sections: true
format:
  html:
    embed-resources: true
    include-in-header: mathjax.html
  pdf: default
  docx: default
---

@@

: Parameters of Maud's statistical model {#tbl-params}

@@

The term $Saturation$ in equation \eqref{eq-decomposition} is a vector of
non-negative real numbers representing, for each reaction, the fraction of
enzyme that is saturated, i.e. bound to one of the reaction's substrates.
To describe saturation we use equation \eqref{eq-saturation}, which is taken
from @liebermeister_modular_2010. Additionally, this term captures competitive
inhibition: as competitive inhibitor concentration increases, the saturation
denominator increases, effectively decreasing the saturation of the substrate on
the total enzyme pool. Conversely, as the substrate concentration increases this
term approaches 1.

@@

: Parameter specification, marginal prior distributions and true parameter values used in our case study. {#tbl-case-study-params}

>>>>

---
title: Supporting information - Bayesian regression facilitates quantitative modelling of cell metabolism
format:
  html:
    embed-resources: true
    include-in-header: mathjax.html
  acs-pdf: 
    include-in-header: 
      - text: \renewcommand{\thepage}{S\arabic{page}}
    keep-tex: true
    journal:
      id: asbcd6  #This is an identifier for the target journal: from https://pubs.acs.org/, the identifier is the part of the URL after https://pubs.acs.org/journal/, e.g. for J. Org. Chem. it would be joceah.
      type: article  #article, note, communication, review, letter, or perspectivee
      layout: traditional #traditional or twocolumn, twocolumn currently broken for longtables
  docx: default
number-sections: true
author: 
  - name: "Teddy Groves"
    orcid: 0000-0002-7109-3270
    email: tedgro@biosustain.dtu.dk
    affiliations: 
      - id: biosustain
        name: DTU
        department: The Novo Nordisk Foundation Center for Biosustainability
        city: Kongens Lyngby 2800
        country: Denmark
    attributes:
      corresponding: true
      equal-contributor: true
    note: These authors contributed equally to this research.
  - name: "Nicholas Luke Cowie"
    orcid: 0000-0003-3367-2105
    affiliations: 
      - ref: biosustain
    attributes:
      equal-contributor: true
    note: These authors contributed equally to this research.
  - name: "Lars Keld Nielsen"
    orcid: 0000-0001-8191-3511
    affiliations: 
      - id: uq
        name: The University of Queensland
        department: Australian Institute for Bioengineering and Nanotechnology (AIBN)
        city: St Lucia 4067
        country: Australia
      - ref: biosustain
bibliography: bibliography.bib
---

{{< pagebreak >}}

@@

: Table S1 – Parameters of Maud's statistical model {#tbl-params}

@@

The term $Saturation$ in equation \eqref{eq-decomposition} is a vector of
non-negative real numbers representing, for each reaction, the fraction of
enzyme that is saturated, i.e. bound to one of the reaction's substrates.
To describe saturation we use equation \eqref{eq-saturation}, which is taken
from @liebermeister_modular_2010 and @noor_note_2013. Additionally, this term captures competitive
inhibition: as competitive inhibitor concentration increases, the saturation
denominator increases, effectively decreasing the saturation of the substrate on
the total enzyme pool. Conversely, as the substrate concentration increases this
term approaches 1.

@@

: Table S2 – Parameter specification, marginal prior distributions and true parameter values used in our case study. {#tbl-case-study-params}
